Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-02-26 write-ups/FSE/paper.tex
Space saving.

The threads of a kernel are divided into \emph{workgroups}, each consisting of a fixed number of threads.  Functions
$\getlocalid$ and $\getgroupid$ return a thread's local id within
its workgroup and the id of the workgroup.
%
%\footnote{For simplicity we assume 1D kernels, which captures all the
%irregular algorithms we know of, and write
%e.g.\ $\getglobalid$ for $\getglobalid(0)$.
%}
%
The number
of threads per workgroup and number of workgroups are obtained via
$\getlocalsize$ and $\getnumgroups$.
%A thread's local and global
%id are related by $\getglobalid = \getgroupid \times
%\getlocalsize + \getlocalid$.  The total number of threads executing the
%kernel is given by $\getglobalsize = \getlocalsize \times \getnumgroups$.

@@

Communication within a workgroup can be achieved
using a workgroup barrier.  Finer-grained
communication within a workgroup, as well as inter-workgroup
communication and communication with the host while the kernel is
running, is enabled by a set of atomic data types and operations.  In
particular, fine-grained host/device communication is via atomic
operations on SVM.

@@

@@

@@

\myparagraph{Work stealing example}
%
Work stealing enables dynamic balancing of tasks across several
processing units. It is useful when the number of tasks to be
processed is dynamic, due to one task creating an arbitrary number of
new tasks.  Work stealing has been explored in the context of
GPUs~\cite{DBLP:conf/egh/CedermanT08,TPO10}. Each workgroup has an
associated queue from which it obtains tasks to process, and to which
it stores new tasks. If a workgroup's queue is empty, the workgroup
tries to \emph{steal} a task from another queue.

\myfiglong\ref{fig:workstealing} shows a simplified version of a work
stealing kernel. Each thread receives a pointer to the
task queues, in global memory, initialized by the
host to contain the initial tasks. Each thread identifies
its workgroup via $\getgroupid$
(line~\ref{line:wksteal:getgroupid}), which is used as a queue id to access the relevant task queue. The
$\mathsf{pop\_or\_steal}$ function (line~\ref{line:wksteal:poporsteal})
pops a task from the workgroup's queue or, if the queue is empty,
tries to steal a task from other queues.

% is called by all threads of a workgroup.  Inside
% $\mathsf{pop\_or\_steal}$ only the thread with a local id equal to $0$,
% the \emph{master thread}, tries to pop a
% task from the workgroup's queue, or to steal a task from a different
% queue. The master thread uses local memory inside
% $\mathsf{pop\_or\_steal}$ to ensure that the returned value is valid for
% all threads of its workgroup.

If a task was obtained, then the workgroup processes it
(line~\ref{line:wksteal:processtask}): during this operation, new tasks
may be created and pushed into the workgroup's queue.

\myparagraph{Graph traversal example} \myfiglong\ref{fig:graphsearch} shows the
skeleton of a frontier-based graph traversal algorithm; such algorithms have
been shown to execute efficiently on GPUs~\cite{BNP12,DBLP:conf/oopsla/PaiP16}.
The kernel is
given three arguments in global memory: a graph structure, and two
arrays of graph nodes. Initially, $\keyword{n0}$ contains the
starting nodes to process. Private variable $\keyword{level}$ indicates the frontier level currently being
processed, and $\keyword{in\_nodes}$ and $\keyword{out\_nodes}$ point to
distinct node arrays recording the nodes to be processed during the current and next frontier, respectively.

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between
threads through \emph{stride} based processing.
%
In this case, the stride is the total number of threads, obtained via
$\getglobalsize$.  Each thread processes nodes with the following sequence of indices:
\code{$\keyword{tid}$+$\keyword{stride}$*0},
\code{$\keyword{tid}$+$\keyword{stride}$*1},
\code{$\keyword{tid}$+$\keyword{stride}$*2}, etc., where $\keyword{tid}$ is a thread's global id.
A thread processes a node via $\keyword{process\_node}$, which may (a) push nodes to process in the next frontier to
$\keyword{out\_nodes}$ and (b) use the current frontier, $\keyword{level}$, in
the computation. After processing the frontier, the threads swap their
node array pointers (line~\ref{line:graph:swap}).

>>>>

The threads of a kernel are divided into \emph{workgroups}.
%, each consisting of a fixed number of threads.
Functions
$\getlocalid$ and $\getgroupid$ return a thread's local id within
its workgroup and the workgroup id.
%
%\footnote{For simplicity we assume 1D kernels, which captures all the
%irregular algorithms we know of, and write
%e.g.\ $\getglobalid$ for $\getglobalid(0)$.
%}
%
The number
of threads per workgroup and number of workgroups are obtained via
$\getlocalsize$ and $\getnumgroups$.
%A thread's local and global
%id are related by $\getglobalid = \getgroupid \times
%\getlocalsize + \getlocalid$.  The total number of threads executing the
%kernel is given by $\getglobalsize = \getlocalsize \times \getnumgroups$.

@@

%Communication within a workgroup can be achieved
%using a workgroup barrier.
Fine-grained
communication within a workgroup, as well as inter-workgroup
communication and communication with the host while the kernel is
running, is enabled by a set of atomic data types and operations.  In
particular, fine-grained host/device communication is via atomic
operations on SVM.

@@

\begin{figure}[t]

@@

\myparagraph{Work stealing}
%
Work stealing enables dynamic balancing of tasks across
processing units. It is useful when the number of tasks to be
processed is dynamic, due to one task creating an arbitrary number of
new tasks.  Work stealing has been explored in the context of
GPUs~\cite{DBLP:conf/egh/CedermanT08,TPO10}. Each workgroup has a
queue from which it obtains tasks to process, and to which
it stores new tasks. If its queue is empty, a workgroup
tries to \emph{steal} a task from another queue.

\myfiglong\ref{fig:workstealing} illustrates a work
stealing kernel. Each thread receives a pointer to the
task queues, in global memory, initialized by the
host to contain the initial tasks. A thread uses its workgroup id 
(line~\ref{line:wksteal:getgroupid}) as a queue id to access the relevant task queue. The
$\mathsf{pop\_or\_steal}$ function (line~\ref{line:wksteal:poporsteal})
pops a task from the workgroup's queue or, if the queue is empty,
tries to steal a task from other queues.
%
% is called by all threads of a workgroup.  Inside
% $\mathsf{pop\_or\_steal}$ only the thread with a local id equal to $0$,
% the \emph{master thread}, tries to pop a
% task from the workgroup's queue, or to steal a task from a different
% queue. The master thread uses local memory inside
% $\mathsf{pop\_or\_steal}$ to ensure that the returned value is valid for
% all threads of its workgroup.
%
If a task was obtained, then the workgroup processes it
(line~\ref{line:wksteal:processtask}), which may lead to new tasks
being created and pushed to the workgroup's queue.

\begin{figure}[t]

@@

\myparagraph{Graph traversal} \myfiglong\ref{fig:graphsearch} illustrates a frontier-based graph traversal algorithm; such algorithms have
been shown to execute efficiently on GPUs~\cite{BNP12,DBLP:conf/oopsla/PaiP16}.
The kernel is
given three arguments in global memory: a graph structure, and two
arrays of graph nodes. Initially, $\keyword{n0}$ contains the
starting nodes to process. Private variable $\keyword{level}$ records the current frontier level, and $\keyword{in\_nodes}$ and $\keyword{out\_nodes}$ point to
distinct arrays recording the nodes to be processed during the current and next frontier, respectively.

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between
threads through \emph{stride} based processing.
%
In this case, the stride is the total number of threads, obtained via
$\getglobalsize$.  Each thread processes nodes with indices
\code{$\keyword{tid}$+$\keyword{stride}$*0},
\code{$\keyword{tid}$+$\keyword{stride}$*1},
\code{$\keyword{tid}$+$\keyword{stride}$*2}, etc., where $\keyword{tid}$ is a thread's global id.
A thread calls $\keyword{process\_node}$ to process a node given the current level, with nodes to be processed next frontier being pushed to $\keyword{out\_nodes}$. After processing the frontier, the threads swap their
node array pointers (line~\ref{line:graph:swap}).
